---
title: "Week 1: SOC 367S"
author: "Andrew Weatherman"
date: 'September 8, 2022'
resources: 
  - "problems/week_1.html"
format: 
  html:
    self-contained: true
    code-tools: true
    highlight-style: github
editor: visual
---

### [Assignment Questions](problems/week_1.Rmd)

### Preparation:

**Load the data:**

```{r, message=FALSE}
load(url('https://dssoc.github.io/datasets/congress.RData'))
load(url('https://dssoc.github.io/datasets/committees.RData'))
```

## Question 1:

The `mean` function returns the average of the passed object. There are three named arguments: `x`, an object, is the R object to be evaluated against; to calculate a symmetrically-trimmed mean, a value \[0, 0.5\] is passed to `trim`, a double, which cuts off that percent of observations from both ends; `na.rm`, a logical, determines whether `NA` values are to be removed before the calculation.

## Question 2:

```{r}
my_vector <- c("R","is","great")
my_vector[3]
```

## Question 3:

```{r}
some_vector <- c(
  25555, 342343, 123123123, 4234234, 53243234, 54324234, 5421111, 12312312, 111231,
  1231231, 12312312, 12312312, 123123, 898972, 789872, 2343, 23423423, 2343221, 23423, 14444, 44324222, 2342341, 124231111, 22233345, 1111233333, 1231231, 1231231
)
```

#### Part A:

```{r}
class(some_vector)
```

#### Part B:

```{r}
max(some_vector)
```

#### Part C:

```{r}
mean(some_vector)
```

#### Part D:

```{r}
54324222 %in% some_vector
```

## Question 4:

```{r}
str(congress)
```

The `congress` object has 539 rows with eight columns and is a `data.frame`.

## Question 5:

We can use `lubridate` to calculate the precise ages of Sherrod Brown and Dianne Feinstein. The `period` class occasionally causes issues down stream, but for this one-off example, it is fine to use.

```{r}
congress |>
  dplyr::filter(
    full_name %in% c(
      "Sherrod Brown",
      "Dianne Feinstein"
    )
  ) |>
  dplyr::mutate(age = lubridate::as.period(
    lubridate::interval(
      birthdate,
      Sys.Date()
    )
  )) |>
  dplyr::select(full_name, age)
```

## Question 6:

For simplicity, I rounded the age to the nearest integer.

```{r}
congress |>
  dplyr::summarize(avg_birthyear = round(mean(birthyear), 0))
```

## Question 7:

While I spend (probably) too much time in R, most of my work is focused in the sports sphere. I have experience with most of what the class covers -- including the `tidyverse`, visualization (with `ggplot2` and `gt/reactable`), web scraping (with `rvest` and `RSelenium`), working with and building APIs, writing and deploying packages to CRAN ([toRvik](www.torvik.dev/)), and some other stuff.

I am taking this course for a chance to apply my skills to a different, yet equally exciting, domain. Without the class, I would probably continue to focus on sports analytics, and while that is really fun, working with different data in different scopes often leads to new ways of thinking and approaching problems.
